---
title: "Reproducible boxplot"
author: "Raoul R. Wadhwa, Chandruganesh Rasendran, Zoran B. Popovic, Steven E. Nissen, Milind Y. Desai"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

This Rmarkdown document shows the code that was used to generate the boxplot in [Wadhwa et al. _JAMA Network Open_. 2021; 4(7): e2118263](https://jamanetwork.com/journals/jamanetworkopen/fullarticle/2782170).
By sharing the R code within this document and uploading the HTML output generated on our local machine, we can make the figure more reproducible without exposing the underlying dataset, which we do not have permission to share.
This refers to the Retraction Watch dataset, which can be acquired via a request at the [Retraction Watch website](https://retractionwatch.com/).

## Code and figure

The code below assumes that the Retraction Watch dataset is saved in the working directory and is named `retractions.csv`.
If the Retraction Watch format is preserved, the block below will generate the boxplot displayed in the [original paper](https://jamanetwork.com/journals/jamanetworkopen/fullarticle/2782170), albeit with formatting differences.

```{r boxplot, fig.width=8, fig.height=4}
# load libraries
library(tidyverse)
library(lubridate)
library(cowplot)

# utility functions
get_decade_bin <- function(y_val) {
  y_val <- as.numeric(y_val)
  vapply(X = y_val,
         FUN.VALUE = character(1),
         USE.NAMES = FALSE,
         FUN = function(year_val) {
           if (year_val <= 1980) {
             "pre-1980"
           } else if (year_val <= 1985) {
             "1981-1985"
           #} else if (year_val <= 90) {
           #  "1986-1990"
             # two five year intervals combined b/c only 2 data points in each before
           } else if (year_val <= 1995) {
             "1986-1995"
           } else if (year_val <= 2000) {
             "1996-2000"
           } else if (year_val <= 2005) {
             "2001-2005"
           } else if (year_val <= 2010) {
             "2006-2010"
           } else {
             "2011-2015"
           }
         })
}

remove_time <- function(val) {
  # remove everything after space
  vapply(X = val,
         FUN.VALUE = character(1),
         FUN = function(curr) {
           if (is.na(curr)) {
             return(NA_character_)
           }
           
           space_index <- as.numeric(regexpr(pattern = " ",
                                             text = curr,
                                             fixed = TRUE))[1]
           
           return(substr(curr, 1, space_index - 1))
         }) -> ans
  
  names(ans) <- NULL
  return(ans)
}

# input data
retract <- read_csv("retractions.csv",
                    col_types = "cccccffccfccccccfcfc")

# manually fill in missing value
retract[retract$`Record ID` == "24122", ]$OriginalPaperDate <- "3/6/18 0:00"

# get info for cardiology articles
cardio_retract <- retract %>%
  # select only medicine-related articles
  filter(grepl(pattern = "(HSC) Medicine - Cardiology",
               x = retract$Subject,
               fixed = TRUE)) %>%
  # convert date columns to Date class
  mutate(RetractionDate = as.Date(remove_time(RetractionDate),
                                  format = "%m/%d/%Y"),
         OriginalPaperDate = as.Date(remove_time(OriginalPaperDate),
                                     format = "%m/%d/%Y"),
         CatchTime = as.numeric(RetractionDate - OriginalPaperDate))

# restructure to get follow-up time trends
cardio_fu_trend <- cardio_retract %>%
  select(RetractionDate, OriginalPaperDate) %>%
  mutate(RetractionDate = lubridate::as_date(RetractionDate),
         OriginalPaperDate = lubridate::as_date(OriginalPaperDate)) %>%
  # calculate days between publication and retraction
  mutate(Delay = as.numeric(RetractionDate - OriginalPaperDate) / 365) %>%
  # remove ones retracted after more than 10 years
  filter(Delay < 5) %>%
  # remove ones w/ less than 10 years follow up (2011 or after)
  filter(year(OriginalPaperDate) < 2016 | year(OriginalPaperDate) > 2020) %>%
  # split into bins by decade of publication
  mutate(Decade = get_decade_bin(year(OriginalPaperDate))) %>%
  # correctly order factor levels
  mutate(Decade = factor(Decade, levels = c("pre-1980", "1981-1985", "1986-1995", "1996-2000", "2001-2005", "2006-2010", "2011-2015")))

# generate boxplot
ggplot(cardio_fu_trend, aes(x = Decade, y = Delay)) +
  geom_boxplot() +
  xlab("Original Publication Date Interval") +
  ylab("Time to Retraction (years)") +
  theme_cowplot()
```
